Complementary roles of Theta and Alpha in Multi-Item Associations

### 2019-12-16 supplemental.tex
chaning to apa style
revisions with replication

\documentclass[11pt, titlepage, twoside]{article}
\usepackage[top=20mm, bottom=20mm, left=20mm, right=20mm]{geometry}
\geometry{a4paper}
\geometry{portrait}
\usepackage[utf8]{inputenc}	% Comment out for xelatex; must use with pdflatex
\usepackage[T1]{fontenc}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{fontspec}
\usepackage{epstopdf}
\usepackage{float}
\usepackage[hang]{footmisc}
\usepackage{endnotes}
\usepackage{listingsutf8}
%\restylefloat{table}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{tabu}
\usepackage{caption}
\usepackage{textcomp}
\usepackage{textgreek}

\usepackage{subcaption}\usepackage[style=numeric,backend=bibtex8]{biblatex}
\addbibresource{references.bib}
\usepackage{longtable}
\usepackage{csvsimple}				% for reading cvs files into tables

@@

@@

The `Deep LEABRA' framework currently being developed in the Emergent software is the initial explorations into the modeling of interactions between thalamic relay nuclei, and superficial vs. deep cortical layers \cite{OReillyWyatteRohrlich14}.  This framework proposes that the neocortex is composed of two separable networks: superficial and deep/thalamic. The superficial-layer network consists of neocortical layers 4, 2, and 3, across different brain areas.  The deep/thalamic network starts in each cortical area with the layer 5b intrinsic bursting (IB) neurons, which receive inputs from local superficial neurons and top-down projections from other areas, and provide strong driving feedforward input to higher-order thalamic nuclei. These 5IB neurons also project to deep layer 6 cortical-thalmic (6CT) neurons, which interconnects with the thalamus (which in turn projects back up to layer 4 of the superficial network and layer 6 in the deep network).  This connectivity is illustrated in Figure \ref{fig:attn_compute}.  The current work takes advantage of the biological referents and computational implementation of the Deep LEABRA frameworkm however, here it is used as a simple form of attentional modulation (and not a temporal learning mechanism as described in \textcite{OReillyWyatteRohrlich14}).  The minimal description of the components necessary for this implementation are described below.

The modulation of standard LEABRA unit activations (referred to as superficial units here) based the Deep LEABRA framework and the thalamic gating model described in \textcite{KetzJensenOReilly15} is carried out in two basic steps.  First, a Thalamic Relay Cell (TRC) unit receives input based on the threshold filtered output of some upstream units which have received its input from processing a given stimulus (in this case the hippocampus is providing this top-down modulation based on what is being retrieved).  These sending units, for a given TRC, are assumed to be the cortical thalamic projections from neocortical layers 5IB, 6CT \cite{ShermanGuillery06}, which reflect a thresholded filtering of superficial layer activity from both top-down and bottom-up sources. TRC activation dynamics are calculated in a similar manner to all LEABRA units, described in detail in \cite{KetzMorkondaOReilly13}, with the exception that its inputs are subject to an activation threshold of 0.1, and the weighted connections into and out of the TRC are not learned but instead remain fixed at 0.8.

Second, the TRC activation is used to calculate the attentional modulation that it sends to down stream superficial units.  The attentional modulation is implemented through a separate set of activations for receiving units that can be considered an attentional mask that gets applied.  This mask is the TRC activation normalized to a range between some pre-determined minimum and a maximum of 1.  The effective modulation of the receiving unit is simply the multiplication of this attentional mask, and the receiving units' net input.  Thus, where activations are strong in the mask, the corresponding superficial layer activations will remain strong, but where they are weaker, the superficial layer activations will be reduced.  This modulation process is hypothesized to be carried out by layer 6CT neurons  \cite{BortoneOlsenScanziani14,OlsenBortoneAdesnikEtAl12}.

This multiplicitive modulation is based off previous mathematical models of attention that capture many of the effects relevant to the perception literature \cite{ReynoldsHeeger09,MontijnKlinkVanWezel12}.  Figure \ref{fig:attn_compute} shows how this framework captures the essential computations of the \textcite{ReynoldsHeeger09} model in different parts of the superficial and deep layer circuits.  Computationally, the TRC activations encode the attentional modulations of the superficial layer state. The attentional modulation signals cause the iterative constraint satisfaction process in the superficial network to focus on task-relevant information while down-regulating responses to irrelevant information, consistent with the abstract \textcite{ReynoldsHeeger09} mode.  Its important to note that this process can work in both a top-down (shown in \ref{fig:attn_compute}) and bottom-up fashion. Biologically, the layer 6CT neurons are known to exhibit a multiplicative influence over firing of superficial-layer neurons, in a manner consistent with the \textcite{ReynoldsHeeger09} model \cite{BortoneOlsenScanziani14,OlsenBortoneAdesnikEtAl12}.

\begin{figure}
  \centering\includegraphics[width=.75\textwidth]{figs/fig_deepleabra_attn_compute.pdf}
  \caption[Signal Flow Through Cortical Layers in Deep LEABRA]{\small How attentional modulation is computed in response to a top-down attentional focus.  Layer 4 receives bottom-up sensory input, which then drives superficial layers (2/3), which initially do not reflect the attentional modulation (not shown).  The deep layer 5 Intrinsic Bursting (5IB) neurons integrate deep-to-deep top-down attentional inputs plus the local stimulus features from 2/3, to produce a {\em raw} deep output, prior to the contextual normalization process.  Layer 6 Cortical-Thalamic (6CT) neurons then integrate this direct activation from 5IB, to produce, for the first time in the circuit, a properly renormalized multiplicative gain-field activation pattern, with surround inhibition both within the 6CT layer and further downstream in the  TRC circuit providing the renormalization process.  These 6CT activations then modulate (multiply) the superficial-layer activations to produce {\em both} an increase in the attended location, and a decrease for the unattended location, as shown.  
  %In biological systems, this modulation also affects the layer 4 inputs (not shown) as well as 2/3, however, our model subsumes layer 4 into layer 2/3 neurons.
  }
  \label{fig:attn_compute}
\end{figure}

>>>>

% !BIB TS-program = biber
\documentclass[11pt, titlepage, twoside]{article}
\usepackage[top=20mm, bottom=20mm, left=20mm, right=20mm]{geometry}
\geometry{a4paper}
\geometry{portrait}
\usepackage[utf8]{inputenc}	% Comment out for xelatex; must use with pdflatex
\usepackage[T1]{fontenc}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{fontspec}
\usepackage{epstopdf}
\usepackage{float}
\usepackage[hang]{footmisc}
\usepackage{endnotes}
\usepackage{listingsutf8}
%\restylefloat{table}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{tabu}
\usepackage{caption}
\usepackage{textcomp}
\usepackage{textgreek}

\usepackage{subcaption}

\usepackage[american]{babel}
\usepackage{csquotes}
\usepackage[style=apa]{biblatex}
\DeclareLanguageMapping{american}{american-apa}

\addbibresource{references.bib}
\usepackage{longtable}
\usepackage{csvsimple}				% for reading cvs files into tables

@@

\usepackage{xr}
\externaldocument{manuscript}

@@

The `Deep LEABRA' framework currently being developed in the Emergent software is the initial explorations into the modeling of interactions between thalamic relay nuclei, and superficial vs. deep cortical layers \parencite{OReillyWyatteRohrlich14}.  This framework proposes that the neocortex is composed of two separable networks: superficial and deep/thalamic. The superficial-layer network consists of neocortical layers 4, 2, and 3, across different brain areas.  The deep/thalamic network starts in each cortical area with the layer 5b intrinsic bursting (IB) neurons, which receive inputs from local superficial neurons and top-down projections from other areas, and provide strong driving feedforward input to higher-order thalamic nuclei. These 5IB neurons also project to deep layer 6 cortical-thalmic (6CT) neurons, which interconnects with the thalamus (which in turn projects back up to layer 4 of the superficial network and layer 6 in the deep network).  This connectivity is illustrated in Figure \ref{fig:attn_compute}.  The current work takes advantage of the biological referents and computational implementation of the Deep LEABRA frameworkm however, here it is used as a simple form of attentional modulation (and not a temporal learning mechanism as described in \textcite{OReillyWyatteRohrlich14}).  The minimal description of the components necessary for this implementation are described below.

The modulation of standard LEABRA unit activations (referred to as superficial units here) based on the Deep LEABRA framework and the thalamic gating model described in \textcite{KetzJensenOReilly15} is carried out in two basic steps.  First, a Thalamic Relay Cell (TRC) unit receives input based on the threshold filtered output of some upstream units which have received its input from processing a given stimulus (in this case the hippocampus is providing this top-down modulation based on what is being retrieved and sent to entorhinal cortex).  These sending units, for a given TRC, are assumed to be the cortical thalamic projections from neocortical layers 5IB, 6CT \parencite{ShermanGuillery06}, which reflect a thresholded filtering of superficial layer activity from both top-down and bottom-up sources. TRC activation dynamics are calculated in a similar manner to all LEABRA units, described in detail in \parencite{KetzMorkondaOReilly13}, with the exception that its inputs are subject to an activation threshold of 0.1, and the weighted connections into and out of the TRC are not learned but instead remain fixed at 0.8.

Second, the TRC activation is used to calculate the attentional modulation that it sends to down stream superficial units.  The attentional modulation is implemented through a separate set of activations for receiving units that can be considered an attentional mask that gets applied.  This mask is the TRC activation normalized to a range between some pre-determined minimum and a maximum of 1.  The effective modulation of the receiving unit is simply the multiplication of this attentional mask, and the receiving units' net input.  Thus, where activations are strong in the mask, the corresponding superficial layer activations will remain strong, but where they are weaker, the superficial layer activations will be reduced.  This modulation process is hypothesized to be carried out by layer 6CT neurons  \parencite{BortoneOlsenScanziani14,OlsenBortoneAdesnikEtAl12}.

This multiplicitive modulation is based off previous mathematical models of attention that capture many of the effects relevant to the perception literature \parencite{ReynoldsHeeger09,MontijnKlinkVanWezel12}.  Figure \ref{fig:attn_compute} shows how this framework captures the essential computations of the \textcite{ReynoldsHeeger09} model in different parts of the superficial and deep layer circuits.  Computationally, the TRC activations encode the attentional modulations of the superficial layer state. The attentional modulation signals cause the iterative constraint satisfaction process in the superficial network to focus on task-relevant information while down-regulating responses to irrelevant information, consistent with the abstract \textcite{ReynoldsHeeger09} model.  Its important to note that this process can work in both a top-down (shown in \ref{fig:attn_compute}) and bottom-up fashion. Biologically, the layer 6CT neurons are known to exhibit a multiplicative influence over firing of superficial-layer neurons, in a manner consistent with the \textcite{ReynoldsHeeger09} model \parencite{BortoneOlsenScanziani14,OlsenBortoneAdesnikEtAl12}.

In this work the minimum TRC activation was set to 0.25 through a loose process of trying to minimize the number of training epochs and matching the pattern of human behavioral performance.  This process consisted of iteratively training networks each with progressively higher minimum TRC modulation values, and finding the network that produced the best paired associate training accuracy with the smallest number of training epochs.   As stated this was loosely fit and a more precise value could be arrived at with more systematic methods.

\begin{figure}
  \centering\includegraphics[width=1\textwidth]{figs/fig_deepleabra_attn_compute_edit.pdf}
  \caption[Signal Flow Through Cortical Layers in Deep LEABRA]{\small How attentional modulation is computed in response to a top-down attentional focus.  \emph{(A)} Layer 4 receives bottom-up sensory input, which then drives superficial layers (2/3), which initially do not reflect the attentional modulation (not shown).  The deep layer 5 Intrinsic Bursting (5IB) neurons integrate deep-to-deep top-down attentional inputs plus the local stimulus features from 2/3, to produce a {\em raw} deep output, prior to the contextual normalization process.  Layer 6 Cortical-Thalamic (6CT) neurons then integrate this direct activation from 5IB, to produce, for the first time in the circuit, a properly renormalized multiplicative gain-field activation pattern, with surround inhibition both within the 6CT layer and further downstream in the  TRC circuit providing the renormalization process.  These 6CT activations then modulate (multiply) the superficial-layer activations to produce {\em both} an increase in the attended location, and a decrease for the unattended location, as shown.  
\emph{(B)} Implementation of computational principle within the Theta-Phase network.  In the model layer 4 is subsumed into layer 2/3 neurons, as represented by ECin and ECout.  Top down influence comes from the hippocampus via area CA1 connections to ECout.  Layer 5IB \emph{raw} signals are calculated from ECin and ECout, and in turn used to calculate the \emph{mod} signal.  This layer 6CT project to the TRC units which ultimately are used to modulate the EC superficial layers with the current attentional mask. 
  }
  \label{fig:attn_compute}
\end{figure}
